---
title: "Shrinkage analysis of fMRI data"
author: "Mengyin Lu"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`


## Data description and visualization
This is a 3D dataset. 21 images each of dimension 64x64.
Each image is a so-called Axial slice i.e imagine salami slicing your brain
from top of your head to the bottom. Each slice is 6mm thick.

This is the result of an fMRI experiment on a single person. While in the
scanner the person was stimulated (visual and auditory stimulation). A time
series of length 180 was collected at each location in the brain and then a
regression is performed at each location, using the stimulation patterns as
regressors. 3 other regressors were used which can be thought of as
confounders. The resulting test statistic is F with df1=2 and df2=175.

```{r}
source("../code/nullalthist.R")
library("oro.nifti")
a = readNIfTI("../data/fstat1.nii.gz")
image(a, col = rainbow(15))
```

## Distribution of the F-statistics
Under the null hypothesis (no signal), the F-statistics should follow F(2, 175). From the histogram, we see that the F-statistics don't quite follow the null distribution F(2, 175), which suggests the exsitence of some signals.
```{r}
hist(a[a !=0], n =10000, xlim = c(0, 20), freq = F, main="Histogram of F-stats")
curve(df(x,df1=2, df2=175), from = 0, to = 10, add = T, col = 2)
legend("topright",col=2,lty=1,legend="F(2,175)")
```

We compute the p-values and use the q-value procedure to control false discovery rate:
```{r}
library(qvalue)
pp = 1-pf(a[a!=0], df1=2, df2=175)
qq = qvalue(pp)
hist(pp, prob=TRUE, main="histogram of p-values")

# estimated null proportion
qvalue(pp)$pi0

# proportion of F-stats declared "significant" with threshold qval<0.05
mean(qq$qvalues<0.05)

# qvalue's implicit partition into null and alternative
lfdr.qv = qq$lfdr
altnullhist(pp,lfdr.qv,main="p values: qvalue",ncz=40,xlab="p value",cex.axis=0.8)

nullalthist(a[a!=0],lfdr.qv,
            main="qvalue's implicit partition of F-stat \n into null and alternative",
            ncz=1000,cex.axis=0.8,xlim=c(0,20),xlab="F stat")
```

The estimated null proportion by the q-value procedure is over 40%.

## Use ASH to shrink the F-stats
Suppose the F statistics follow the distribution $F_j/\alpha_j \sim F(2,175)$. Under the null, $\alpha_j=1$. Our shrinkage procedure assumes that 
\[\log(\alpha_j) \sim g(\cdot),\]
where $g(\cdot)$ is an unimodal distribution with mode at 0.

We use the empirical Bayes procedure to estimate $g$, and use the posterior mean of $\log(\alpha_j)$ as a shrunk estimate of $\log(alpha_j)$. 

```{r}
library(ashr)
test = ash(log(a[a!=0]),1,lik=logF_lik(df1=2,df2=175),mixcompdist="+uniform")

# replace the raw signals by the shrunk signals
a.shrink = a
a.shrink[a!=0] = exp(test$result$PosteriorMean)

# visualize data after shrinkage
image(a.shrink, col = rainbow(15), main="after FASH shrinkage")
```

The estimated $g$ and null proportion by ASH are:
```{r}
# estimated g
test$fitted_g

# estimated null proportion
test$fitted_g$pi[1]

# proportion of F-stats declared "significant" with threshold qval<0.05
mean(test$result$qvalue<0.05)

# ASH's implicit partition into null and alternative
lfdr.ash = test$result$lfdr
nullalthist(a[a!=0],lfdr.ash,
            main="ASH's implicit partition of F-stat \n into null and alternative",
            ncz=1000,cex.axis=0.8,xlim=c(0,20),xlab="F stat")
```

## Session Information

```{r session-info}
```
